An efficient composition of bidirectional programs by tupling and lazy updates

### 2019-11-27 paper/1.introduction.tex
add bib

%In software, there are strong demands for synchronizing data. In database community this is known as ``the view update problem'' and researched for a long time \cite{viewupdate}.
% A Survey to View Update Problem 
%As a solution for this problem, bidirectional transformation (BX) is introduced.
The synchronization of data is a common problem. In the database community this problem is known as “the view update problem” and has been investigated for a long time~\cite{viewupdate}. Bidirectional transformation (BX) is an approach to solve this problem.
As an example, let us consider a small BX program of $phead$\footnote{The actual program is shown in the next section.}. BX provides two functions: $get$ and $put$. Figure \ref{fig:eval-phead} denotes the behaviors of these functions when evaluating $phead$.

@@

First, let us see the standard evaluation method of compositions: ``not keeping any intermediate states and obtaining them by evaluation when they are needed.'' In this example, ``intermediate states'' are \texttt{i$_1$} and \texttt{i$_2$}. This method is used in a BX language, BiGUL \cite{josh_pepm} \cite{josh_popl}. The merit of this method is clear semantics and easy to implement. The disadvantage is the number of $get$s will be quadratic when the BX programs' compositions are left associative. In the evaluation of lp3, two $get$s are required for obtaining \texttt{i$_2$} and one $get$ is required for obtaining \texttt{i$_1$}. In total, three $get$s are evaluated. We explain this method in Section 2.

>>>>

%In software, there are strong demands for synchronizing data. In database community this is known as ``the view update problem'' and researched for a long time.
% A Survey to View Update Problem 
%As a solution for this problem, bidirectional transformation (BX) is introduced.
The synchronization of data is a common problem. In the database community this problem is known as “the view update problem” and has been investigated for a long time~\cite{Bancilhon:1981:USR:319628.319634}. Bidirectional transformation (BX) is an approach to solve this problem.
As an example, let us consider a small BX program of $phead$\footnote{The actual program is shown in the next section.}. BX provides two functions: $get$ and $put$. Figure \ref{fig:eval-phead} denotes the behaviors of these functions when evaluating $phead$.

@@

First, let us see the standard evaluation method of compositions: ``not keeping any intermediate states and obtaining them by evaluation when they are needed.'' In this example, ``intermediate states'' are \texttt{i$_1$} and \texttt{i$_2$}. This method is used in a BX language, BiGUL \cite{Ko:2016:BFV:2847538.2847544,Ko:2017:ABB:3177123.3158129}. The merit of this method is clear semantics and easy to implement. The disadvantage is the number of $get$s will be quadratic when the BX programs' compositions are left associative. In the evaluation of lp3, two $get$s are required for obtaining \texttt{i$_2$} and one $get$ is required for obtaining \texttt{i$_1$}. In total, three $get$s are evaluated. We explain this method in Section 2.

### 2019-11-27 paper/6.related-work.tex
add bib

There are many BX languages \cite{josh_pepm} \cite{BXtend} \cite{eMoflon} \cite{EVL+Strace} \cite{JTL} \cite{NMF}
%(memo: BXtend, eMoflon, EVL+Strace, JTL, NMF. SDMLib should be added?)
based on several application scenarios and BX lenses works \cite{} \textcolor{red}{(add papers for lenses)}. The semantics and correctness of bidirectional programs has been investigated intensively during the past years, but their efficiency and optimization are not yet fully understood. Anjorin et. al. introduces the first benchmark for BX languages and compared them \cite{BXcomp}. However, the improvement for practical implementation of BX languages is still missing. As a first step, we study different evaluation methods to optimized their evaluation. We can say this paper will be first attempt to improve efficiency of BX composition evaluation. In this paper we focus on a BX language BiGUL \cite{josh_pepm} \cite{josh_popl}, we compare BiGUL's method (in Sect. 2) and our methods (in Sect. 3 and 4). From experimental results of left associative BX composition programs, we know that our memoization methods are faster than the original BiGUL's evaluation method.
In this paper we focus on BiGUL, however our methods can be applicable to other BX languages which have similar problems.

@@

Memoization \cite{} is well-known technique to reduce extra evaluation. In our case, simple application of memoization $put_m$ works well for many BX programs.
Another way to reduce extra evaluation is fusion \cite{}. However, fusion of BX programs are not considered well \textcolor{red}{(really?)}. Therefore we did not use this way.
  Instead of that, we introduce another way that uses tupling \cite{tupling}, lazy update and lazy computation \cite{}. Tupling is often used for reducing common evaluations, but we use it for reducing the size of complements.
  Lazy evaluation is also a well-known and widly-used optimization technique to reduce the redundant evaluation. Because we use OCaml for implementation, we need to introduce this explicitly. If we use Haskell, a lazy evaluated language, we do not need to introduce them.

>>>>

There are many BX languages \cite{Buchmann:2018:BFI:3362232.3362263,Ko:2016:BFV:2847538.2847544,LeblebiciAS14,Samimi-Dehkordi18,Cicchetti2011,Hinkel:2019:CPB:3318595.3318634}
%(memo: BXtend, eMoflon, EVL+Strace, JTL, NMF. SDMLib should be added?)
based on several application scenarios and BX lenses works \cite{} \textcolor{red}{(add papers for lenses)}. The semantics and correctness of bidirectional programs has been investigated intensively during the past years, but their efficiency and optimization are not yet fully understood. Anjorin et. al. introduces the first benchmark for BX languages and compared them \cite{Anjorin2019}. However, the improvement for practical implementation of BX languages is still missing. As a first step, we study different evaluation methods to optimized their evaluation. We can say this paper will be first attempt to improve efficiency of BX composition evaluation. In this paper we focus on a BX language BiGUL \cite{Ko:2016:BFV:2847538.2847544, Ko:2017:ABB:3177123.3158129}, we compare BiGUL's method (in Section 2) and our methods (in Section 3 and 4). From experimental results of left associative BX composition programs, we know that our memoization methods are faster than the original BiGUL's evaluation method.
In this paper we focus on BiGUL, however our methods can be applicable to other BX languages which have similar problems.

@@

Memoization \cite{} is well-known technique to reduce extra evaluation. In our case, simple application of memoization $put_m$ works well for many BX programs.
Another way to reduce extra evaluation is fusion \cite{}. However, fusion of BX programs are not considered well \textcolor{red}{(really?)}. Therefore we did not use this way.
  Instead of that, we introduce another way that uses tupling \cite{Fokkinga90}, lazy update and lazy computation \cite{}. Tupling is often used for reducing common evaluations, but we use it for reducing the size of complements.
  Lazy evaluation is also a well-known and widly-used optimization technique to reduce the redundant evaluation. Because we use OCaml for implementation, we need to introduce this explicitly. If we use Haskell, a lazy evaluated language, we do not need to introduce them.

### 2019-11-27 paper/main.tex
small fixes
change on TODO part
add bib

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage{graphicx}
\usepackage{color}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{tabularx}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{pgfplotstable}
\usepackage{lmodern}
\usepackage{pgffor}
\usepackage{schemata}

% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}
\usetikzlibrary{matrix}
\usepgfplotslibrary{groupplots}
\pgfplotsset{compat=newest}

\setlength{\parindent}{0pt}
\renewcommand{\indent}{\hspace*{0pt}}

@@

\authorrunning{}

@@

\institute{anonymous}

@@

\textcolor{red}{TODO: FIX reference, order, contents ..}

% From sample code
% 
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
% \bibliographystyle{splncs04}
% \bibliography{mybibliography}
%

\tiny

\begin{thebibliography}{8}
\bibitem{viewupdate}
F. Bancilhon and N. Spyratos. ``Update semantics of relational views''. ACM Transactions on Database Systems, 6(4):557--575, 1981.

\bibitem{BXcomp}
Anthony Anjorin, Thomas Buchmann, Bernhard Westfechtel, Zinovy Diskin, Hsiang-Shang Ko, Romina Eramo, Georg Hinkel, Leila Samimi-Dehkordi, and Albert Zündorf
``Benchmarking bidirectional transformations: Theory, implementation, application, and assessment'', Software and Systems Modeling, ??, 2019

\bibitem{tupling}
  M. Fokkinga. ``Tupling and mutumorphisms''. Squiggolist, 1(4), 1989.

\bibitem{josh_pepm}
  Hsiang-Shang Ko, Tao Zan, Zhenjiang Hu, BiGUL: A Formally Verified Core Language for Putback-Based Bidirectional Programming, ACM SIGPLAN 2016 Workshop on Partial Evaluation and Program Manipulation (PEPM 2016), pp.61--72, 2016.

\bibitem{josh_popl}
  Hsiang-Shang Ko, Zhenjiang Hu, An Axiomatic Basis for Bidirectional Programming , 45th ACM SIGPLAN Symposium on Principles of Programming Languages (POPL 2018), 2018.

\bibitem{BXtend}
  Thomas Buchmann. BXtend -- a framework for (bidirectional) model transformations. In Slimane Hamoudi,Luis Ferreira Pires, and Bran Selic, editors, Proceedings
of the 6th International Conference on Model-Driven Engineering and Software Development - Volume 1: MODELSWARD (MODELSWARD 2018), pages 336--345, 2018.

\bibitem{eMoflon}
  Erhan Leblebici, Anthony Anjorin, and Andy Sch"{u}rr. Developing eMoflon with eMoflon. In Davide Di Ruscio and Daniel Varro, editors, Theory and Practice of Model Transformations - 7th International Conference, ICMT 2014, Held as Part of STAF 2014, York, UK, 2014. Proceedings, volume 8568 of Lecture Notes in
Computer Science, pages 138–145. Springer, 2014.

\bibitem{EVL+Strace}
Leila Samimi-Dehkordi, Bahman Zamani, and Shekoufeh Kolahdouz-Rahimi. EVL+Strace: A novel bidirectional transformation approach. Information and Software
Technology, 100:47--72, 2018.

\bibitem{JTL}
Antonio Cicchetti, Davide Di Ruscio, Romina Eramo, and Alfonso Pierantonio. JTL: A bidirectional and change propagating transformation language. In Brian Malloy, Steffen Staab, and Mark van den Brand, editors, Proceedings of the Third International Conference on Software Language Engineering (SLE 2010), volume 6563 of Lecture Notes of Computer Science, pages 183--202, Springer-Verlag.

\bibitem{NMF}
Georg Hinkel and Erik Burger. Change propagation and bidirectionality in internal transformation DSLs. Soft-ware and Systems Modeling, 18(1):249--278, 2019.

\bibitem{SDMLib}
  SDMLib paper?

\bibitem{lenses}
  Lense papers?

\bibitem{tutorial}
  Tutorial paper of BiGUL

@@

%\bibitem{ref_url1}
%LNCS Homepage, \url{http://www.springer.com/lncs}. Last accessed 4
%Oct 2017
\end{thebibliography}

>>>>

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage{graphicx}
\usepackage{color}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{tabularx}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{pgfplotstable}
\usepackage{lmodern}
\usepackage{pgffor}
\usepackage{schemata}
\usepackage[colorlinks=true,linkcolor=blue,citecolor=blue]{hyperref}

\usepackage[backend=bibtex,doi=false,url=false,style=nature,sorting=none]{biblatex}
\addbibresource{bibliography}

% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
\renewcommand\UrlFont{\color{blue}\rmfamily}

\usetikzlibrary{matrix}
\usepgfplotslibrary{groupplots}
\pgfplotsset{compat=newest}

% \setlength{\parindent}{0pt}
% \renewcommand{\indent}{\hspace*{0pt}}

@@

\authorrunning{anonymous authors}

@@

%\institute{anonymous}

@@

\textcolor{red}{TODO: fix references (order, contents), add missing references}

% From sample code
% 
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
% \bibliography{bibliography}
% \bibliographystyle{splncs04}

\printbibliography

% \tiny

% \begin{thebibliography}{8}
% \bibitem{viewupdate}
% F. Bancilhon and N. Spyratos. ``Update semantics of relational views''. ACM Transactions on Database Systems, 6(4):557--575, 1981.

% \bibitem{BXcomp}
% Anthony Anjorin, Thomas Buchmann, Bernhard Westfechtel, Zinovy Diskin, Hsiang-Shang Ko, Romina Eramo, Georg Hinkel, Leila Samimi-Dehkordi, and Albert Zündorf
% ``Benchmarking bidirectional transformations: Theory, implementation, application, and assessment'', Software and Systems Modeling, ??, 2019

% \bibitem{tupling}
%   M. Fokkinga. ``Tupling and mutumorphisms''. Squiggolist, 1(4), 1989.

% \bibitem{josh_pepm}
%   Hsiang-Shang Ko, Tao Zan, Zhenjiang Hu, BiGUL: A Formally Verified Core Language for Putback-Based Bidirectional Programming, ACM SIGPLAN 2016 Workshop on Partial Evaluation and Program Manipulation (PEPM 2016), pp.61--72, 2016.

% \bibitem{josh_popl}
%   Hsiang-Shang Ko, Zhenjiang Hu, An Axiomatic Basis for Bidirectional Programming , 45th ACM SIGPLAN Symposium on Principles of Programming Languages (POPL 2018), 2018.

% \bibitem{BXtend}
%   Thomas Buchmann. BXtend -- a framework for (bidirectional) model transformations. In Slimane Hamoudi,Luis Ferreira Pires, and Bran Selic, editors, Proceedings
% of the 6th International Conference on Model-Driven Engineering and Software Development - Volume 1: MODELSWARD (MODELSWARD 2018), pages 336--345, 2018.

% \bibitem{eMoflon}
%   Erhan Leblebici, Anthony Anjorin, and Andy Sch"{u}rr. Developing eMoflon with eMoflon. In Davide Di Ruscio and Daniel Varro, editors, Theory and Practice of Model Transformations - 7th International Conference, ICMT 2014, Held as Part of STAF 2014, York, UK, 2014. Proceedings, volume 8568 of Lecture Notes in
% Computer Science, pages 138–145. Springer, 2014.

% \bibitem{EVL+Strace}
% Leila Samimi-Dehkordi, Bahman Zamani, and Shekoufeh Kolahdouz-Rahimi. EVL+Strace: A novel bidirectional transformation approach. Information and Software
% Technology, 100:47--72, 2018.

% \bibitem{JTL}
% Antonio Cicchetti, Davide Di Ruscio, Romina Eramo, and Alfonso Pierantonio. JTL: A bidirectional and change propagating transformation language. In Brian Malloy, Steffen Staab, and Mark van den Brand, editors, Proceedings of the Third International Conference on Software Language Engineering (SLE 2010), volume 6563 of Lecture Notes of Computer Science, pages 183--202, Springer-Verlag.

% \bibitem{NMF}
% Georg Hinkel and Erik Burger. Change propagation and bidirectionality in internal transformation DSLs. Soft-ware and Systems Modeling, 18(1):249--278, 2019.

% \bibitem{SDMLib}
%   SDMLib paper?

% \bibitem{lenses}
%   Lense papers?

% \bibitem{tutorial}
%   Tutorial paper of BiGUL

% \bibitem{putget}
%   H. Pacheco, Z. Hu, and S. Fischer. Monadic combinators for “putback” style bidirectional programming. In Partial Evaluation and Program Manipulation, pages 39–50. ACM, 2014.

% \bibitem{putput}

@@

%\bibitem{ref_url1}
%LNCS Homepage, \url{http://www.springer.com/lncs}. Last accessed 4
%Oct 2017
% \end{thebibliography}
